---
title: "Regular Expression - Examples"
author: "Asmik & Lisa - for Intro to NLP"
date: "30 April 2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r setup_packages, message=FALSE}
# Load required packages

library("stringr")
```

### Disclaimer

With regular expressions there are often several ways to specify the desired pattern.
In the following, we will have a look at the most frequently used text mining operations, employing various regex patterns.

You are free to use whatever works but it's certainly good practice to pick a style and stick to it as it makes your code more easily readable.

In any case do not hesitate to ask Google. Thanks to regex being present in many programming languages there is plenty of help out there.

The [`stringr` vignette](https://cran.r-project.org/web/packages/stringr/vignettes/regular-expressions.html) provides a nice intro.

### Identify

The most generic command to match regular patterns is probably `str_detect`, returning a boolean value indicating whether or not a string contains the specified pattern.

Identify a **specific character sequence**:

```{r identify_0}
stringr::str_detect(
  string = c("Hello world", "bye"), 
  pattern = "world")
```

Identify **any character** value:

```{r identify_1}

string <- c("999_user_999", "123")

list(
  posix = stringr::str_detect(string, pattern = "[:alpha:]"),
  ascii = stringr::str_detect(string, pattern = "[A-Za-z]"))
```

Identify **lowercase** characters:

```{r identify_2}

string <- c("why ON EARTH", "WHAT!!?!1!")

list(
  posix = stringr::str_detect(string, pattern = "[:lower:]"), # [:upper:] for uppercase
  ascii = stringr::str_detect(string, pattern = "[a-z]")) # [A-Z] for uppercase
```

Identify **digits**:

```{r identify_3}

string <- c("I got 99 problems but regex ain't one", "one two three")

list(
  perl = stringr::str_detect(string, pattern = "\\d"),
  posix = stringr::str_detect(string, pattern = "[:digit:]"),
  ascii = stringr::str_detect(string, pattern = "[0-9]"))
```

Identify **punctuation**:

```{r identify_4}

stringr::str_detect(
  string = c("Well... you're right!", "No punct to be seen here"), 
  pattern = "[:punct:]")
```

Identify **spaces**:

```{r identify_5}

string <- c("Gimme some space, will ya", "tightstuffhere")

list(
  perl = stringr::str_detect(string, pattern = "\\s"),
  posix = stringr::str_detect(string, pattern = "[:space:]"))
```

Caution with special characters -- we can still query those but we need to "escape" them by prefixing them with a backslash. This tells R that we mean a literal symbol here.

Even more caution with characters that have a special meaning in regex. For example, `+` is used to indicate an occurrence of at least one repetition (more on this in a moment). Internally, to give the plus sign its special meaning in the first place, regex prefix them by a backslash also. In order to match `+`, we therefore need *double* backslashes.

Identify **special characters**:

```{r identify_6}

stringr::str_detect(
  string = "Cote d'Azure", 
  pattern = "\'") # omitting the backslash will throw an error
```

```{r identify_7}
stringr::str_detect(
  string = "Wanna be my +1?", 
  pattern = "\\+")
```

Now let's talk **repetitions**.

Basically, we have the following possibilities:

* `?` repeat 0x or 1x
* `*` repeat 0x or more
* `+` repeat 1x or more
* `{n}` repeat exactly n times
* `{m,n}` repeat between m and n times
* `{m,}` repeat at least m times
* `{,n}` repeat at most n times

Matches 3 p's in a row:

```{r identify_9}

stringr::str_detect(
  string  = c("nlp_2021", "nlppp_2021"),
  pattern = "p{3}")
```

Matches 2 dots in a row:

```{r identify_10}

stringr::str_detect(
  string  = c("never.stop.learning", "never stop learning.."),
  pattern = "\\.{2}")
```

Matches at most one l:

```{r identify_11}
stringr::str_detect(
  string = c("np", "nlp", "nllp", "nllllp"),
  pattern = "nl?p")
```

Matches at least two l's:

```{r identify_12}
stringr::str_detect(
  string = c("np", "nlp", "nllp", "nllllp"),
  pattern = "nl{2,}p")
```

Matches between 1 and 2 l's:

```{r identify_13}
stringr::str_detect(
  string = c("np", "nlp", "nllp", "nllllp"),
  pattern = "nl{1,2}p")
```

Matches one or more quotation marks:

```{r identify_14}
stringr::str_detect(
  string = c("What?", "What??", "Huh"),
  pattern = "\\?+")
```

### Extract

`str_extract` works similarly but returns the specified string instead of `TRUE` or `FALSE`.

We will extract character sequences and have a look at **groups** wrapped in parentheses.

For example, we can extract sequences with some optional parts, which might be helpful in dealing with plurals:

```{r extract_1}
stringr::str_extract(
  string = c("Mark my word", "words don't do her justice"),
  pattern = "word(s)?")
```

We can also use grouping to specify various accepted options separated by `|`:

```{r extract_2}
stringr::str_extract(
  string = c("Mum, is there any pudding left?", "No mom, I want fries!", "Mommy please"),
  pattern = "(M|m)(u|o)m(my)?")
```

### Locate

Besides detecting and extracting patterns, we sometimes need to know their exact location within a text.
`str_locate` returns the beginning and end of the specified pattern (`NA` if the pattern is not detected).

Make sure to allow for repetitions if needed:

```{r locate_1}

stringr::str_locate(
  string = c("nlp2021", "nlp_kurs"),
  pattern = "\\d")
```

```{r locate_2}

stringr::str_locate(
  string = c("nlp2021", "nlp_kurs"),
  pattern = "\\d+")
```

Sometimes we might expect multiple occurrences of our patterns.
In that case we use `str_locate_all` (`str_locate` only returns the location of the first match -- the same is true for, e.g., `str_replace` and `str_replace_all`).
When in doubt, better go for `_all`.

```{r locate_3}
stringr::str_locate_all(
  string = c("I am Thorin son of Thrain son of Thror, King under the Mountain!"),
  pattern = "of")
```

### Replace

Text cleaning will often require replacement of unwanted, leading or trailing sequences.

`^` and `$` are **anchors** indicating the beginning and end of a string, respectively.

Replace leading 0's by an empty character (we could also use `str_remove_all` here): 

```{r replace_1}

stringr::str_replace(
  string = c("0005", "1050"),
  pattern = "^0+",
  replacement = "")

```

Replace gender suffices by "x": 

```{r replace_2}
stringr::str_replace(
  string = c("Innenarchitektin", "Binnenschifffahrtskapitänin", "Professorinnen"),
  pattern = "in(nen)?$",
  replacement = "x")
```

Besides the simple `^` and `$` there is another useful anchoring pattern, namely **look-arounds**. 
This allows to match a pattern only if it is preceded or followed by a specific other pattern:

* Look-ahead `(?=...)`: match if followed by "..."
* Look-behind `(?<=...)`: match if preceded by "..."
* Negated version for both: replace `=` by `!`

Replace any word preceded by `#` with "hashtag":

```{r replace_3}
stringr::str_replace_all(
  string = c("Ob Laschet Kanzler wird? #idautit"),
  pattern = "(?<=#)[:alpha:]+",
  replacement = "hashtag")
```

Remove all numbers except when they are followed by a currency expression:

```{r replace_4}
stringr::str_replace_all(
  string = c("I bought 5 apples for 5$"),
  pattern = "\\d+(?!\\$)",
  replacement = "")
```

Or, simply replace occurrences at any location (make sure to use the `_all` version): 

```{r replace_5}
stringr::str_replace_all(
  string = c("Our Father, who art in heaven, hallowed be thy name; thy kingdom come, thy will be done on earth as it is in heaven."),
  pattern = "thy",
  replacement = "your")
```

### Other useful operations

`stringr` offers a lot more than we have covered so far. 

Some additional operations that might be useful:

```{r misc_1}
stringr::str_split(
  string = c("Heufer-Umlauf", "Walter-Borjans"),
  pattern = "-")
```

```{r misc_2}
stringr::str_count(
  string = c("lorem ipsum, quia dolor sit"),
  pattern = "i")
```

```{r misc_3}
stringr::str_squish(
  string = c("   if  you manipulate   strings this sometimes leaves  annoying whitespaces   "))
```


**As you might have realized by now, handling regex is often simply trial-and-error (especially when you try to match more complex sequences). They will most certainly drive you crazy at some point or other, but they are a powerful and absolutely essential tool in text mining.**